---
title: "SHIN-Enzyme vs NA"
author: "qh"
date: "2/7/2019"
output: html_document
---
#Load all packages
```{r}
library(RSepc)
library(ggplot2)
library(pracma)
library(psych)
library(base)
library(baseline)
library(tidyverse)
library(stats)
```

#read data
```{r}
setwd("/Users/qinghe/Desktop/R/2019/1-30 gpa au-enzyme shin/1-30-2019-pga")
filenames <- list.files(path = "/Users/qinghe/Desktop/R/2019/1-30 gpa au-enzyme shin/1-30-2019-pga")
nfile <- length(filenames)
sample<-do.call("cbind", lapply(filenames, read.csv, header = FALSE))
nrow(sample)
t = c(1:(2*nfile))
t = t[c(FALSE,TRUE)]
t = c(1,t)
sample.new=sample[,t]
SN <- sample.new
colnames(SN)<-c("WtNo",c(1:8))
#write.table(SN,"/Users/qinghe/Desktop/R/onoin au enzyme/r files/onoinauenzyme.csv",
#            row.names=FALSE,col.names=FALSE,sep=",")
#COMBINE DATA IN CSV FILE
FE <- sample.new
```

#formate
```{r}
#RESHAPE
#after combining, data is then reshaped into the long format
setwd("/Users/qinghe/Desktop/R/2019/1-30 gpa au-enzyme shin/1-30-2019-pga")

#eliminate file type info ".CSV"
FN <- unlist(strsplit(filenames, "[.]"))
FN1 <- FN[seq(1,2*length(filenames),2)]
FN.1 <- unlist(strsplit(FN1, "[-]"))

#long form this no/group/sample format can be used by Rsepc
colnames(FE)<-c("WtNo",FN1)
library(reshape)
fe.long<-melt(FE[,2:(nfile + 1)])
fe.long$variable <- as.character(fe.long$variable)
x <- unlist(strsplit(fe.long$variable, "[-]"))
fe.long$class<-x[seq(3,nrow(sample)*4*length(filenames),4)]
length(filenames)
group<-x[seq(1,4*length(filenames),4)]
group <- unlist(strsplit(colnames(FE)[-1], "[-]"))
groups <- NA
groups <- group[seq(3, 4*length(filenames),4)]
fe.long$rep<-x[seq(4, nrow(sample)*4*length(filenames),4)]
fe.long$WtNo<-NA
fe.long$WtNo <- rep(FE$WtNo,length(filenames))
length(fe.long$WtNo)
colnames(fe.long)<-c("spec.no","Intensity","group","sample","WtNo")
```

#preprocessing
```{r}
#BASELINE CORRECTION
sp_poly <- baseline_corrt(fe.long, degree = 13, tol = 1e-3, rep = 100)
#baseline corrrection methods see baseline
Intensity <- t(sp_poly)
fe.long$Intensity<-Intensity[,2]
fe <- head(fe.long,3528)#nrow(sample)
plot(fe.long$WtNo,fe.long$Intensity,type="l", pch=20,cex =0.1)

#SMOOTHING
s.fe.long <- smooth_sp(fe.long)
# check smoothing result
fe <- head(fe.long,3528)#nrow(sample)
plot(fe$WtNo,fe$Intensity,type="l")
s.fe <- head(s.fe.long,3528)
plot(s.fe$WtNo,s.fe$Intensity,type="l")
```

#Spectrum difference before normalization
The sample without nanoparticles show higher intensity. This may indicate that the SERS inert species may not be able to be enhanced by SHIN.
```{r}
averager1 <- average_sp(s.fe.long , gname = "shin")
averager2 <- average_sp(s.fe.long , gname = "na")
avepgar <- cbind(averager1,averager2$`Average Intensity`)
colnames(avepgar) <- c("WtNo","shin","na")

matplot(avepgar$WtNo,avepgar[,c(2,3)],type="l",col = 1:2,lty=1,xlab="WtNo", ylab= "shin/na")
legend(100,100, legend=c("shin","na"),col=1:2, lty=1, cex=0.8)
```

#NORMALIZATION
In the following section, the average spectrum for sample PG(substrate) with Au-Enzyme and SHIN mixture and pure PG substrate are calculated. Note that the pure substrate sample without nanoparticles are labeled as NA indicating the absence of nanoparticles.

The t test has been run for each raman shift number. The t test result had been converted to binnary form, all P-value larger than 0.05 are shown as 0, indicating there is no siginifcant difference between 2 groups, and all P-value smaller than 0.05 are shown as 1, indicating siginificant difference between 2 groups. 

The average spectrum and t test result are shown in the same plot to illustrate the spectrum difference. 

##intensity norm + test
Normalize based on highest intensity peak. 
The significant peak difference can be located around 1600 cm-1 (may due to the 4ATP contamination), the peak around 2800 cm-1.
```{r}
#NORMALIZATION
#intensity
sp_intensity<-intensitynorm_sp(s.fe.long)
sp_wide_deer <- spread(sp_intensity,WtNo,Intensity)

p <- ttest_sp(sp_intensity)
tdata <- subset(p, (X1 == "shin")& (X2 == "na"))
tdata$value[tdata$value >= 0.05] <- 1
tdata$value[tdata$value < 0.05] <- 0

average1 <- average_sp(sp_intensity, gname = "shin")
average2 <- average_sp(sp_intensity, gname = "na")

std1 <- std_sp(sp_intensity, gname = "shin")
std2 <- std_sp(sp_intensity, gname = "na")

avepga <- cbind(average1,average2$`Average Intensity`)
colnames(avepga) <- c("WtNo","shin","na")


matplot(avepga$WtNo,avepga[,c(2,3)],type="l",col = 1:2,lty=1,xlab="WtNo", ylab= "shin/na")
#matplot(avepga$WtNo,avepga[,c(2,3)],type="l",col = 1:2,lty=1,xlab="WtNo", ylab= "shin/na")
legend(100,0.8, legend=c("shin","na"),col=1:2, lty=1, cex=0.8)
lines(tdata$wtno,(1-tdata$value)*0.8,type = "l", col = 3)
```

#NORMALIZATION
##peak norm + test
Normalize the spectrum with the peak area from 800 to 900 cm-1.
The significant peak difference can be located around 1600 cm-1 (may due to the 4ATP contamination), the peak around 2800 cm-1
```{r}
#peak
sp_peaknorm <- peaknorm_sp(s.fe.long, peak.range = c(800,900))
sp_wide_deer2 <- spread(sp_peaknorm,WtNo,Intensity)


pp <- ttest_sp(sp_peaknorm)
tdatap <- subset(pp, (X1 == "shin")& (X2 == "na"))
tdatap$value[tdatap$value >= 0.05] <- 1
tdatap$value[tdatap$value < 0.05] <- 0

averagep1 <- average_sp(sp_peaknorm, gname = "shin")
averagep2 <- average_sp(sp_peaknorm, gname = "na")

stdp1 <- std_sp(sp_peaknorm, gname = "shin")
stdp2 <- std_sp(sp_peaknorm, gname = "na")

avepgap <- cbind(averagep1,averagep2$`Average Intensity`)
colnames(avepgap) <- c("WtNo","shin","na")


matplot(avepgap$WtNo,avepga[,c(2,3)],type="l",col = 1:2,lty=1,xlab="WtNo", ylab= "shin/na")
#matplot(avepga$WtNo,avepga[,c(2,3)],type="l",col = 1:2,lty=1,xlab="WtNo", ylab= "shin/na")
legend(100,0.8, legend=c("shin","na"),col=1:2, lty=1, cex=0.8)
lines(tdatap$wtno,(1-tdatap$value)*0.8,type = "l", col = 4)

```

#NORMALIZATION
##area norm + test
Normalize based on the total peak area. This is not a good way for normalization.
```{r}
#area
sp_areanorm <- areanorm_sp(s.fe.long) 
sp_wide_deer3 <- spread(sp_areanorm,WtNo,Intensity)

pa <- ttest_sp(sp_areanorm)
tdataa <- subset(pa, (X1 == "shin")& (X2 == "na"))
tdataa$value[tdataa$value >= 0.05] <- 1
tdataa$value[tdataa$value < 0.05] <- 0

averagea1 <- average_sp(sp_areanorm, gname = "shin")
averagea2 <- average_sp(sp_areanorm, gname = "na")

stda1 <- std_sp(sp_areanorm, gname = "shin")
stda2 <- std_sp(sp_areanorm, gname = "na")

avepgaa <- cbind(averagea1,averagea2$`Average Intensity`)
colnames(avepgaa) <- c("WtNo","shin","na")


matplot(avepgaa$WtNo,avepgaa[,c(2,3)],type="l",col = 1:2,lty=1,xlab="WtNo", ylab= "shin/na")
#matplot(avepga$WtNo,avepga[,c(2,3)],type="l",col = 1:2,lty=1,xlab="WtNo", ylab= "shin/na")
legend(100,8, legend=c("shin","na"),col=1:2, lty=1, cex=0.8)
lines(tdataa$wtno,(1-tdataa$value)*6,type = "l", col = 5)
#lines(tdata$wtno,tdata$value*6,type = "l", col = 4)
legend(1000,8, legend=c("area","intensity"),col=3:4, lty=1, cex=0.8)

```

